Social Housing in Toronto

### 2024-04-21 paper/paper.qmd
commit

library(tidyverse)
library(palmerpenguins)
```

@@

@@

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

@@

# Extract relevant columns with adjusted column names
year_quarter <- raw_data$Quarter_Year
new_reactivated_applications <- raw_data$New_reactivated
housed <- raw_data$Housed

# Create a data frame
plot_data <- data.frame(Year_Quarter = year_quarter,
                        New_Reactivated_Applications = new_reactivated_applications,
                        Housed = housed)

# Create a line plot
ggplot(plot_data, aes(x = Year_Quarter)) +
  geom_line(aes(y = New_Reactivated_Applications, color = "New/Reactivated Applications"), size = 1.2) +
  geom_line(aes(y = Housed, color = "Housed"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Yearly New/Reactivated Applications vs. Housed",
       x = "Year Quarter",
       y = "Application Status") +
  theme_minimal() +
  theme(legend.title = element_blank())

@@

@@

# Extract relevant columns with adjusted column names
quarter/year <- raw_data$Quarter/Year
total_active <- raw_data$Total_active_waiting_list
seniors <- raw_data$Seniors

# Create a data frame
plot_data <- data.frame(Quarter/Year = quarter/year,
                        Total_active_waiting_list = total_active,
                        Seniors = seniors)

# Convert Quarter/Year to factor to maintain order
plot_data$Quarter/Year <- factor(plot_data$Quarter/Year, levels = unique(plot_data$Year/Quarter))

# Create a line plot
ggplot(plot_data, aes(x = Quarter_Year)) +
  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
  geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Total Active Applications vs. Seniors' Applications",
       x = "Quarter Year",
       y = "Application Numbers") +
  theme_minimal() +
  theme(legend.title = element_blank())

@@

@@

>>>>

library(tidyverse)
library(palmerpenguins)

analysis_data <- read_csv(here::here("data/analysis_data.csv"))
```

@@

The data in this post was compiled using open source statistical programming language R (RCoreTeam). It uses functionalists from readxl, dplyr, ggplot2, here, and readr. The cleaning process focuses on transforming the data from the third column onward. It uses the mutate and across functions in dplyr to standardize the responses by converting any values that are not "yes" or "no" to NA (missing values), ensuring that the data set contains only clear binary responses for analysis. This step is critical to maintaining the consistency and reliability of the data set, as it filters out ambiguous or invalid entries. Finally, the cleaned data is saved as a CSV file in the specified "analysis_data" directory. This allows access to the cleaned data set for further analysis or reporting. Using write.csv with line name = FALSE ensures the output file is neat.

@@

To clean the data for better analysis processing, necessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed.

@@

@@

**Figure 1:**

@@

# Extract relevant columns with adjusted column names
year_quarter <- analysis_data$quarter_year
new_reactivated_applications <- analysis_data$new_reactivated
housed <- analysis_data$housed

# Create a data frame
plot_data <- data.frame(
              quarter_year = year_quarter,
              test = new_reactivated_applications,
              housed = housed
            )

# Convert Quarter/Year to factor to maintain order
plot_data$quarter_year <- 
  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))

# Create a plot
ggplot(plot_data, 
       aes(x = quarter_year)) +
  geom_point(aes(y = test), color = "red") +
  geom_point(aes(y = housed), color = "blue") +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Yearly New/Reactivated Applications vs. Housed",
       x = "Year Quarter",
       y = "Application Status") +
  theme_minimal() +
  theme(legend.title = element_blank())

@@

**Figure 2:**

@@

# Extract relevant columns with adjusted column names
year_quarter <- analysis_data$quarter_year
new_reactivated_applications <- analysis_data$new_reactivated
housed <- analysis_data$housed

# Create a data frame
plot_data <- data.frame(
              quarter_year = year_quarter,
              test = new_reactivated_applications,
              housed = housed
            )

# Convert Quarter/Year to factor to maintain order
plot_data$quarter_year <- 
  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))

# Create a line plot
ggplot(plot_data, aes(x = quarter_year)) +
  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
  geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Total Active Applications vs. Seniors' Applications",
       x = "Quarter Year",
       y = "Application Numbers") +
  theme_minimal() +
  theme(legend.title = element_blank())

@@

**Figure 3:**

@@

**Figure 4:**

# Extract relevant columns with adjusted column names
year_quarter <- analysis_data$quarter_year
new_reactivated_applications <- analysis_data$new_reactivated
housed <- analysis_data$housed

# Create a data frame
plot_data <- data.frame(
              quarter_year = year_quarter,
              test = new_reactivated_applications,
              housed = housed
            )

# Convert Quarter/Year to factor to maintain order
plot_data$quarter_year <- 
  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))

# Extract relevant columns
inactive_cancelled <- analysis_data$inactive_cancelled
total_active_waiting_list <- analysis_data$total_active_waiting_list

# Add the new data to the existing plot_data data frame
plot_data$Inactive_Cancelled <- inactive_cancelled
plot_data$Total_Active_Waiting_List <- total_active_waiting_list

# Create a line plot
ggplot(plot_data, aes(x = quarter_year)) +
  geom_line(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled"), size = 1.2) +
  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Inactive/Cancelled vs Total Active Waiting List",
       x = "Quarter Year",
       y = "Application Status") +
  theme_minimal() +
  theme(legend.title = element_blank())
